---
title: "Stats Exam, 31st August 2016"
Author: Luyolo Matutu
output: github_document
---
# Question 1

## Null Hypothesis
* increasing resting heart rate in males does not have a stronger association with increasing body temperature compared with females

## Alternative Hypothesis
* increasing resting heart rate in males has a stronger association with increasing body temperature compared with females

```{r echo = TRUE}

# read the data
x <- read.csv("file:///C:/Users/Wits-User/Desktop/Stats Exam 2016/question1.csv")

#plot the data for male heart rate vs body temperature
plot(x$male,x$body_temperature, main = 'The effect of increasing resting heart rate on body temperature in males', ylab = 'body temperature', xlab = 'male resting heart rate', pch = 19, col = "turquoise")
abline(lm(x$male~x$body_temperature, data = x), lwd = 3, col = "black")

# linear regression
x1 <- lm(x$male~x$body_temperature, data = x)
summary(x1)

#plot the data for female heart rate vs temperature
plot (x$female,x$body_temperature, main = 'The effect of increaing resting heart rate on body tempereature in females', ylab = 'body temperature', xlab = 'female resting heart rate', pch = 19, col = "purple")
abline(lm(x$female~x$body_temperature, data = x), lwd = 3, col = "black")

#linear regression
x1 <- lm(x$female~x$body_temperature, data = x)
summary(x1)
```

## Test assumptions
* linear regression aims to see if there is a linear relationship between the duration of running time and the calories consumped.
* the data has two variables, therefore, one variable is considered to be independent and the other is considered to be a dependent variable.

## Output of statistical test
* Males: df = 63, p value = 0.1333
* Females: df = 63, p value = 0.02186

## Conclusions
* p value for males is more than 0.05 but the p value for females is less than 0.05, therefore reject the alternative hypothesis and accept the null hypothesis.
* females have a stronger association between resting heart rate and body temperature.

# Question 2

## Null hypopthesis
* intoxicated test subjects do not use their dominant hand and are not subject to ataxic gait.

## Alternative hypothesis
* intoxicated individuals use their dominant hand and are subject to ataxic gait.
```{r echo=TRUE}
# read the data
x <- read.csv("file:///C:/Users/Wits-User/Desktop/Stats Exam 2016/question2.csv")
x1 <- xtabs(~handedness + first_stumble, data = x)
head(x)
table(x$handedness, x$first_stumble)

#contents of data x
dim(x)
names(x)
head(x)
tail(x)
summary(x)

#plot the data

barplot(x1, beside = TRUE, ylab = 'intoxication', xlab= 'presentation', main = 'effect of intoxication on handedness and ataxia', col= c("pink", "turquoise"))

legend('top', c("handedness", "ataxia"), fill = c("pink", "turquoise"))

# statistical test
z <- chisq.test(x1, correct = TRUE)
z
```
## Test assumption
* the data is categorical but from the same population
* the aim of the exercise is to see if there is an association between the variables with intoxication

## Output of the statistical test
* df = 1, p value < 0.05

## Conclusion
* Alcohol consumption causes ataxia and more dependence on dominant hand

# Question 3

## Null hypothesis
* The number of calories consumed does not depend on running time

## Alternative hypothesis
* The number of calories consumed depends on running time

```{r echo=TRUE}
# read the data
x <- read.csv("file:///C:/Users/Wits-User/Desktop/Stats Exam 2016/question3.csv")

#tidy the data
time = x$time
calories = x$calories
head(cbind(time,calories))

#plot the data
plot(x$time,x$calories, main = 'the effect of running time on calorie consumption', xlab = 'time (s)', ylab = 'calorie consumption (cal)', pch = 19, col = "pink")
abline(lm(x$time ~ x$calories, data = x), col = "purple", lwd = 3)

#linear regression
x1 <- lm(x$time~x$calories, data = x)
summary(x1)
```

## prediction of calories consumed if running time is reduced to 30 minutes
* 460 calories

## Test assumptions
* linear regression aims to see if there is a linear relationship between the duration of running time and the calories consumped.
* the data has two variables, therefore, one variable is considered to be independent and the other is considered to be a dependent variable.

## Output of the statistical test
* p value: 0.001522, df: 17

## Conclusion
* p value< 0.05 therefore reject the null hypothesis and accept the alternative hypothesis.
* there is a linear relationship between the two variables, therefore calorie consumption depends on the duration of running time.

# Question 4
```{r}

```

